---
title: "assignment"
author: "Jacques Chatard"
date: "April 2 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Purpose of this project  and data source

This project is related to the human activity recognition 'HAR'.
We will use the WLE dataset collected by :   

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative 
Activity Recognition of Weight Lifting Exercises. Proceedings of 4th 
International Conference in Cooperation with SIGCHI (Augmented Human '13) . 
Stuttgart, Germany: ACM SIGCHI, 2013.

[Many thanks for their generosity]

We will use and compare 3 different models on WLE (weight lifting exercises) dataset.
And we will try to estimate the accuracy of the predictions on the  'classe' 
variable which classifies the quality of execution of weight lifting movements.

# Loading and reading data
 - Loading data:

```{r}
destfile="pml-training.csv"
fileURL<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
if (!file.exists(destfile)) {
        download.file(fileURL ,destfile,method="auto")
}

```
```{r}
destfile="pml-testing.csv"
fileURL<- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if (!file.exists(destfile)) {
        download.file(fileURL ,destfile,method="auto")
}
```
- Reading data:
```{r}
pml_training<-read.table(file = "pml-training.csv", header=TRUE, sep=",")
pml_testing <-read.table(file ="pml-testing.csv", header=TRUE, sep=",")
dim(pml_training); dim(pml_testing)

```

# loading the required libraries:
```{r libraries, message=FALSE}
library(caret); library(kernlab); library(randomForest); library(dplyr)
library(parallel); library(foreach); library(doParallel)

```

The original data is already divided into two groups of data: pml-training and pml-testing.
Nevertheless, pml-testing represents only 1 per thousand of the total data.
I would proceed to a new subdivision of the data from pml-training.

# Cleaning the data:  




